### 2018-07-18 Paper/old.tex
..

\documentclass[11pt]{article}

\usepackage{lineno,hyperref}
\usepackage{latexsym,amsthm,amsmath,xcolor,graphicx, amssymb, color,hyperref,booktabs,mathrsfs}

%\modulolinenumbers[5]

\usepackage{geometry}
 \geometry{
 letterpaper,
 total={8.5in,11in},
 left=.75in,
 right=.75in,
 top=.75in,
 bottom=.75in,
 }

% Macros
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}
\newcommand{\norm}[1]{\left\| {#1} \right\|}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\grad}{\nabla}
\newcommand{\hessian}{\nabla^2}
\newcommand{\V}{\mathbb{V}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\ST}{\mathbb{S}^{\text{tot}}}
\newcommand{\ip}[2]{\langle{#1}, {#2}\rangle}
\newcommand{\tp}{{\!\mathsf{T}}}
\newcommand{\trace}{\mathrm{trace}}
\newcommand{\mat}[1]{\mathbf{{#1}}}
\newcommand{\Np}{{N_\text{p}}}
\newcommand{\E}[1]{\mathrm{E}\left\{{#1} \right\}}

\newcommand{\alennote}[1]{{\color{blue} Alen: {#1}}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{proposition}{Proposition}[section]

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
\title{Efficient UQ for chemical systems using active 
\author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan}
subspaces}

\begin{document}
\maketitle

\begin{abstract}
Insert abstract here.
\end{abstract}

\tableofcontents
\section{Introduction}

Constributions:
\begin{enumerate}
\item Iterative computation of active subspaces for efficient surrogate modeling 
\item Comparison of gradient-free and gradient based approaches.
\item Perspectives on activity scores as means of approximating DGSMs.
\item Application to challenging chemical kinetics problem. 
\end{enumerate}
Insert introduction here.

\section{Active subspaces and sensitivity analysis}

\subsection{Active subspaces}
We recall some background regarding active subspaces~\cite{Constantine15}.
\subsection{Global sensitivity measures}
Define the total Sobol' indices $T_i(f)$ and discuss their use.

For $f: \Omega \to \R$, we consider the DGSMs~\cite{SobolKucherenko09},
\[
    \nu_i(f) := \E{\left(\frac{\partial f}{\partial\xi_i}\right)^2} =
                  \int_\Omega 
                  \left(\frac{\partial f}{\partial\xi_i}\right)^2
                  \pi(\vec{\xi})d\vec{\xi}, \quad i = 1, \ldots, \Np.   
\]
Here $\pi$ is the joint PDF of $\xi$. 
Let $\mat{C} = \E{\nabla f \nabla f^T}$ and note that
\[
   \nu_i(f) = \vec{e}_i^T \mat{C} \vec{e}_i, 
\]
where $\vec{e}_i$ is the $i$th coordinate vector in $\R^\Np$.
This also shows that $\sum_{j=1}^\Np \nu_i(f) = \trace(\mat{C})$.
We consider the spectral decomposition of $\mat{C}$, given by  
$\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^T$, where 
$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$ and
$\vec{u}_k$ are the corresponding (orthonormal) eigenvectors.
We note that,
\begin{equation}\label{equ:spectral_DGSM}
\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^T\Big) \vec{e}_i
 = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2, 
\end{equation}
which gives a spectral representation for the DGSMs. Truncating the summation 
leads to the notion of activity scores introduced in~\cite{Diaz16,ConstantineDiaz17}
defined as follows: 
\[
   \alpha_i(f; r) =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2,
   \quad j = 1, \ldots, \Np, \quad r \leq \Np.
\]
The activity scores connect ideas from active subspaces and global sensitivity
analysis, and can be used to approximate DGSMs.  The following result, which
can be found in~\cite{Diaz16,ConstantineDiaz17}, quantifies the error in this
approximation. We provide a short proof for completeness. 
\begin{proposition}\label{prp:dgsm_bound} 
For $1 \leq r \leq \Np$,
\[
0 \leq \nu_i(f) - \alpha_i(f; r) \leq \lambda_{r+1}, \quad i = 1, \ldots, \Np.
\] 
\end{proposition}
\begin{proof}

%Using the spectral representation of the DGSMs and the definition of activity
%scores we clearly see:
%\[
%\alpha_i(f;r) = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 \leq \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 = \nu_i(f), \quad \quad i = 1,\ldots,N_p, \quad r \leq N_p
%\]
%In other words
%\[
%0 \leq  \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 - \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
%\]
%with equality if $N_p=r$.
%\newline
%We can write:
%\[
%\begin{aligned}
%\nu_i(f) = \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2  \\
%= \alpha_i(f;r) + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 \leq \alpha_i(f;r) + \lambda_{r+1} \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{u}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
%\end{aligned}
%\]
%The eigenvectors $\vec{u}_k$ are orthonormal so they all have length 1. Also note that for every $x \in \R^n$ we have \[ \norm{x}^2 = \sum_{k=1}^{n} \langle \vec{x}, \vec{u}_k \rangle^2\] This is known as Parseval's identity.
%In particular in this case $\vec{x} = \vec{e}_i \in \R^{N_p}$ so
%\[1 = \norm{\vec{e}_i}^2 = \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{u}_k \rangle^2\]
%Finally we write:
%\[
%\nu_i(f) \leq \alpha_i(f;r) + \lambda_{r+1}, \quad \quad i = 1,\ldots,N_p
%\]

Note that, $\nu_i(f) - \alpha(f; r)  = \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \geq 0$,
which gives the first inequality. To see the upper bound, we note,
\[
   \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \leq \lambda_{r+1} \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2
   \leq \lambda_{r+1}. 
\]
The last inequality holds because 
$1 = \norm{\vec{e}_i}^2 = 
\sum_{k = 1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2 
\geq \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2$.
\end{proof} 
The utility of this result is realized in problems with 
high-dimensional parameter spaces in which 
the eigenvalues $\lambda_i$ decay rapidly to zero; in 
such cases, this result implies that  $\nu_i(f) \approx \alpha_i(f; r)$,
where $r$ is the \emph{numerical rank} of $\mat{C}$.  This will be especially
effective if there is a large gap in the eigenvalues.

Consider a function $y = f(\xi_1, \xi_2, \ldots, \xi_\Np)$ where
$\xi_i$ are independent are either uniformly distributed or
distributed according to a Boltzman probability distribution.  
It was shown in~\cite{LamboniIoossPopelinEtAl13} that the total Sobol' 
index $T_i(f)$ can be bounded in terms of $\nu_i(f)$:
\begin{equation}\label{equ:sobol_bound}
T_i(f) \leq \frac{C_i}{\V(f)}\nu_i(f), \quad i = 1, \ldots, \Np,
\end{equation}
where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant.
This provides a strong theoretical basis for using DGSMs to identify 
unimportant inputs. Moreover, 
in~\cite{VohraAlexanderianSaftaEtAl18} we introduced the normalized screening metric
\[
   \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
\]
for detecting unimportant input parameters. 
Note that if $\xi_i$ are iid, the $C_i$'s will cancel and the normalized screening
metric reduces to 
\[
    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
\]
The following result is a straightforward generalization of
the result from~\cite{Diaz16}, and 
links activity scores and Sobol' indices.
\begin{proposition}
For $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
\[
       T_i(f) \leq \frac{C_i}{\V(f)}(\alpha_i(f; r) + \lambda_{r+1}).
\]
\end{proposition}
\begin{proof}
The result follows immediately from Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}.
\end{proof}
We can also bound the normalized DGSMs using activity scores; it is easy to see  
\[
\tilde{\nu}_i(f) \leq 
\frac{ C_i \big(\alpha_i(f; r) + \lambda_{r+1}\big)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)}
=\frac{C_i \alpha_i(f; r)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)} + \kappa_i \lambda_{r+1}, 
\]
whith $\kappa_i = C_i / (\sum_i C_i \alpha_i(f; r))$. 
In the case where where $\lambda_{r+1} \approx 0$, 
this motivates definition of
normalized activity scores
\[
   \tilde{\alpha}_i(f; r) =  \frac{C_i \alpha_i(f; r)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)}.
\]

\section{Method}

\section{Application to $H_2$/$O_2$ reaction mechanism: high-dimensional case}

\alennote{Manav, can we consider a higher dimensional problem with uncertain
activation energies as well?}

\section{Conclusion}

\section*{References}

\bibliography{mybibfile}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.40\textwidth]{./Figures/comp_eig}
\hspace{1mm}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_eigv}
  \\ \vspace{5mm}
  \includegraphics[width=0.40\textwidth]{./Figures/comp_ssp}
\hspace{1mm}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \\ \vspace{5mm}
  \includegraphics[width=0.42\textwidth]{./Figures/ub_conv_kinetics_rich}
\end{center}
\end{figure}

\end{document}