### 2018-09-25 Paper/abstract.tex
..

\section*{Abstract}
We focus on an efficient approach for quantification of uncertainty in complex chemical
reaction networks with a large number of uncertain parameters.  Parameter
dimension reduction is accomplished by computing an~\emph{active subspace} that
predominantly captures the variability in the quantity of interest (QoI).  In
the present work, we compute the active subspace for a H$_2$/O$_2$ mechanism 
that involves 19 chemical reactions, using an efficient iterative strategy.  The
active subspace is first computed for a 19-parameter problem wherein only the
uncertainty in the pre-exponents of the individual reaction rates is
considered. This is followed by the analysis of a 33-dimensional case wherein
the activation energies are also considered uncertain.  In both cases, a
1-dimensional active subspace is identified, which indicates enormous
potential for efficient statistical analysis of complex chemical systems. In
addition, we explore links between active subspaces and global sensitivity
analysis, and exploit these links for identification of key
contributors to the variability in the model response.