### 2018-11-29 Revision/Paper/app.tex
..

\section{$\text{H}_2$/$\text{O}_2$ reaction kinetics: higher-dimensional case}
\label{sec:app}

For the high-dimensional case, we aim to investigate the impact of the uncertain
pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) \rebut{and
the initial conditions: pressure~($P_0$), temperature~($T_0$), and stoichiometry based on
the equivalence ratio~($\Phi_0$)} on ignition delay during the H$_2$/O$_2$ reaction. 
\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except 
$\mathcal{R}_6$ -- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to 
zero nominal values for $E_a$), and the initial conditions were perturbed by
2$\%$ about their nominal values. Note that the mangnitude of the perturbation
was selected such that the ignition delay remains positive throughout the input
domain. The nominal values of the rate parameters, $A_i$'s and $E_{a,i}$'s 
were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and
$\Phi_0$ were considered to be 1.0 atm, 900~K, and 2.0 respectively.}

\subsection{Computing the active subspace}

The active subspace was computed using the iterative procedure outlined in 
Algorithm~\ref{alg:grad}. The convergence of the eigenvectors was examined
by tracking the quantity `$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
\rebut{In Figure~\ref{fig:conv_app}, we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
with increasing iterations for the perturbation and the regression approaches 
discussed earlier in Section~\ref{sec:method}}. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/eig_conv36Dp2}
\caption{\rebut{Left: An illustrative comparison of individual components of the 
dominant eigenvector in the converged active subspace i.e., at the end
of 4 iterations in the perturbation approach and 9 iterations in the
regression approach. Right: A comparison of the convergence behavior
of the perturbation and the
regression approaches. Convergence is accomplished once 
$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
than 0.05.}}
\label{fig:conv_app}
\end{center}
\end{figure}
%

In Figure~\ref{fig:hd} (left), we illustrate converged estimates of the eigenvalues.
 The second eigenvalue is
found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right), generated
using both, perturbation-based and regression-based approaches for estimating
the gradient. 
%
\begin{figure}[htbp]
 \begin{center}
   \includegraphics[width=0.45\textwidth]{./Figures/eig_33D}
   \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: SSP of the computed active subspace, and the
corresponding 1-D linear fit.} 
\label{fig:hd}
\end{center}
\end{figure}
%
Specifically
in the case of perturbation-based approach, the SSP is observed to exhibit a linear variation that
is reasonably
captured by the corresponding straight-line fit, $\tilde{G}$, constructed using the sequence of steps
outlined in Section~\ref{sub:ac}.
On the other hand, the SSP obtained by estimating the gradient using regression,
is relatively more scattered. Consequently, the two surrogates (straight-line fits) are observed to exhibit a
discrepancy unlike our earlier observation for the 19-dimensional problem in Figure~\ref{fig:comp_ssp}.

\subsection{Surrogate Assessment}
\label{sub:verify}

The 1-dimensional surrogate ($\tilde{G}$) shown in Figure~\ref{fig:hd}~(right) is investigated
for its accuracy as well as the ability to capture the uncertainty in the
model output in two ways. Firstly, we estimate the relative L$_2$ norm of the discrepancy~($\varepsilon_d$)
between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
model output and the surrogate in the following equation:
%
\be
\varepsilon_d = \frac{\|G(\bm{\xi}) - \tilde{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
\ee
%
Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used to evaluate $\varepsilon_d$. 
It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
using $\tilde{G}$ based on the perturbation-based and the regression-based approaches respectively. Hence,
$\varepsilon_d$ was found to increase by an order of magnitude when the gradient is estimated using
the regression-based approach.
Therefore, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas,
the regression-based approach resulted in a relative error of about 10$\%$.

Secondly, we verified the accuracy of the two surrogates in a probabilistic setting. In particular, we compared 
probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
surrogates ($\tilde{G}$'s) based on the two approaches for gradient estimation as shown in Figure~\ref{fig:pdf_33D}.
Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the cross-validation set. 
%
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3.0in]{./Figures/pdf_comp_id_1e4}
\end{center} 
\caption{A comparison of the PDFs of ignition delay, obtained using model 
evaluations (solid line) and 1-dimensional surrogates using the regression-based strategy (dashed line) and the
perturbation-based strategy (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were 
used in each case.}
\label{fig:pdf_33D}
\end{figure}
%
%
%\begin{figure}[htbp]
% \begin{center}
%  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}
%\caption{A comparison of the PDFs of ignition delay, obtained using model evaluations (solid line) and
%1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based approach
%(dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were used in each case.}
%\label{fig:pdf_33D}
%\end{center}
%\end{figure}
%
The PDFs based on the model evaluations and the perturbation-based approach are nearly identical, confirming
the existence of a 1-dimensional active subspace as well as the accuracy of gradient estimates in this case. 
On the other hand, although the PDF obtained using the regression-based approach  is observed to show
consistency with the true PDF in the modal
estimate, it clearly underestimates the spread in the ignition delay. Specific values of the mean and the standard 
deviation in each case are provided in Table~\ref{tab:stats}. 
%
\begin{table}[htbp]
\begin{center}
\begin{tabular}{ccc}
\toprule
$\textbf{Distribution}$ & $\mu$ & $\sigma$ \\ 
\bottomrule
$G$~(Model) & 0.133 & 0.0198 \\
$\tilde{G}$~(Perturbation-based) & 0.133 & 0.0196 \\
$\tilde{G}$~(Regression-based) & 0.133 & 0.0167 \\
\bottomrule
\end{tabular}
\caption{The mean ($\mu$), and the standard deviation ($\sigma$), computed using the model ($G$), and
the surrogate ($\tilde{G}$) based on the two strategies at 10$^4$ samples in the cross-validation
set.}
\label{tab:stats}
\end{center}
\end{table}
%
The three PDFs are found to be
consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates using model
evaluations and the perturbation-based approach are found to be in close agreement, whereas, it  is accurately
 estimated in the case of regression-based approach only upto the first significant digit.

\subsection{GSA consistency check}

The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace, obtained
using the two approaches for gradient estimation (perturbation-based and regression-based),
are compared with the 
total Sobol' indices
in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
1-dimensional surrogate ($\tilde{G}$) in the active subspace, obtained using the 
perturbation-based approach. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new2}
\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
\label{fig:as_33D}
\end{center}
\end{figure}
%
Several useful inferences can be drawn. Firstly, the total Sobol indices are found to be
consistent
with the normalized activity scores as obtained when the gradient is estimated using a perturbation-based approach.
The regression-based
approach, however, yields consistent results for the $A_i$'s but fails capture the sensitivity associated with one
of the most important parameters, $E_{a,15}$. This is because numerical errors are incurred when approximating
the model gradients using a linear regression fit in this case. 
Secondly, our findings indicate that   
the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
found to be low or negligible. 
%Note that this observation can also be exploited for reducing the 
%dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
%The sensitivity-driven dimension reduction for uncertainty quantification has been
%discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
%subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
%shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.