### 2018-11-30 Revision/Paper/conc.tex

\section{Summary and Discussion}
\label{sec:conc}

In this work, we focused on the uncertainty associated with the
rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism
\rebut{as well as the initial pressure, temperature, and stoichiometry} and its
impact on ignition delay predictions. The mechanism involves 19 different
reactions and in each case, the reaction rate depends upon the choice of a
pre-exponent and an activation energy. Hence, in theory, the evolution of the
chemical system depends upon 38 rate parameters \rebut{and three initial conditions}.
However, we considered 
epistemic uncertainty in all pre-exponents and activation energies with non-zero
nominal values i.e. a total of 33 rate parameters instead of 38 \rebut{in addition to
the three initial conditions}.  
%Conventional means for uncertainty quantification such as those involving
%surrogate model construction as well as sensitivity analysis would thus be
%computationally challenging.
To facilitate efficient uncertainty analysis, we focused our efforts on
reducing the dimensionality of the problem by identifying important directions
in the parameter space such that the model output 
predominantly varies along these directions. These important directions
constitute the active subspace. Additionally, we demonstrated that the activity scores,
computed using the components of the dominant eigenvectors provide an efficient
means for approximating derivative based global sensitivity measures (DGSMs).
Furthermore, we established generalized mathematical linkages between the
different global sensitivity measures: activity scores, DGSMs, and total Sobol'
index which could be exploited to reduce computational effort associated with
global sensitivity analysis.

Active subspace computation requires repeated evaluations of the gradient of
the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
namely, perturbation-based (e.g. finite difference - used in this work, adjoint-based
methods, etc.) and regression-based. Both approaches were shown to yield consistent
results for the 19-dimensional
problem wherein only the pre-exponents were considered to be uncertain.
\rebut{It was observed that the computational effort required to obtain a 
converged active subspace was comparable for the two approaches. However,
the predictive accuracy of the perturbation approach was observed to be
slightly higher. Moreover, a 1-dimensional active subspace was shown to
reasonably approximate the uncertainty in the ignition delay.}
Additionally, the activity scores were also shown to be consistent with the
screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
iterative procedure was adopted to enhance the computational efficiency.

The active subspace was further computed for a 36-dimensional problem
wherein all pre-exponents and activation energies with non-zero nominal
estimates \rebut{as well as the initial conditions} were considered uncertain. 
\rebut{Once again, consistent results were obtained using the two approaches.
A 1-dimensional active subspace was shown to reasonably
capture the uncertainty in the ignition delay in this case. However, the
computational effort required to compute a converged active subspace
using perturbation was found to be half of the effort required in the case
of regression. Predictive accuracy of the two approaches was found to 
be comparable. Hence, perturbation seems like a preferred approach
for the higher-dimensional problem based on our findings. GSA results indicated
that the variability in the ignition delay is predominantly due to the 
uncertainty in the rate parameters, $A_1$ and $A_9$ with significant
contributions from $A_{15}$, $A_{17}$, and $E_{a,15}$. Additionally, the
ignition delay was found to be sensitive towards $T_0$.}

Therefore, in general, perturbation-based approaches should be preferred for 
active subspace computation especially in situations where efficient gradient
computation techniques (adjoints-based, automatic differentiation, etc.) are
feasible. The regression-based approach can be explored in situations involving intensive
simulations where the perturbation-based methods are intractable.
%
%and the goal is obtaining rough estimates of the statistics of the QoI as
%opposed to a more detailed analysis such as those involving parametric
%sensitivity.  
%
We also mention that alternate regression based approaches such as ones based
on computing a global quadratic model have been proposed and used in the
literature; see e.g.,~\cite{Constantine:2017a}.  The applicability of such an
approach in the context of high-dimensional chemical reaction networks is
subject to future work.

The computational framework presented in this work is agnostic to the choice of
the chemical system and can be easily adapted for other systems as long
as the quantity of interest is continuously differentiable in the considered
domain of the inputs.  We have demonstrated that the active subspace could be
exploited for efficient forward propagation of the uncertainty from inputs to
the output. The resulting activity scores and the low-dimensional surrogate
could further guide optimal allocation of computational resources for
calibration of the important rate-controlling parameters \rebut{and input conditions}
 in a Bayesian setting.
Additionally, dimension reduction using active subspaces could assist in
developing robust formulations for predicting discrepancy between simulations
and measurements due to epistemic uncertainty in the model inputs.