### 2019-01-21 KineticsPaperArXiv/ack.tex

\section*{Acknowledgment}

M.~Vohra and S.~Mahadevan gratefully acknowledge funding support from the
National Science Foundation (Grant No. 1404823, CDSE Program), and Sandia
National Laboratories (PO No. 1643376, Technical monitor: Dr. Joshua Mullins).
The research of A.~Alexanderian was partially supported by the
National Science Foundation through the grant DMS-1745654.  The authors also
thank Dr.~Cosmin Safta at Sandia National Laboratories for his guidance
pertaining to the usage of the TChem software package.